### 2019-09-04 initial/notes.tex
check in new canonical coords + eom

\subsection{Canonical Coordinates}

Consider canonical coordinates $x = \sin \frac{\theta}{2}\cos \phi, y = \sin
\frac{\theta}{2}\sin \phi$, then the Hamiltonian can be written
\begin{align}
    H &= -\frac{1}{2}\cos^2\theta
        + \eta \p*{\cos \theta \cos I - \sin I \sin \theta \cos \phi}
        ,\nonumber\\
        &= -\frac{1}{2}\p*{1 - 2\sin^2\frac{\theta}{2}}^2
            + \eta\p*{\p*{1 - 2\sin^2\frac{\theta}{2}}\cos I
            - 2 \sin \frac{\theta}{2}\cos \phi \cos \frac{\theta}{2}\sin I}
            ,\nonumber\\
        &= -\frac{1}{2}\p*{1 - 2\p*{x^2 + y^2}}^2
            + \eta\p*{\p*{1 - 2\p*{x^2 + y^2}}\cos I
                - 2\sin I x\sqrt{1 - \p*{x^2 + y^2}}}.
\end{align}

>>>>

\subsection{Another Set of Canonical Coordinates}

Consider canonical coordinates $x = 2\sin \frac{\theta}{2}\cos \phi, y = 2\sin
\frac{\theta}{2}\sin \phi$ (these can be verified to be canonical by evaluating
PB $\z*{y, x}_{\cos \theta, \phi} = 1$), then the Hamiltonian can be written
\begin{align}
    H &= -\frac{1}{2}\cos^2\theta
        + \eta \p*{\cos \theta \cos I - \sin I \sin \theta \cos \phi}
        ,\nonumber\\
        &= -\frac{1}{2}\p*{1 - 2\sin^2\frac{\theta}{2}}^2
            + \eta\p*{\p*{1 - 2\sin^2\frac{\theta}{2}}\cos I
            - 2 \sin \frac{\theta}{2}\cos \phi \cos \frac{\theta}{2}\sin I}
            ,\nonumber\\
        &= -\frac{1}{2}\p*{1 - \frac{x^2 + y^2}{2}}^2
            + \eta\p*{\p*{1 - \frac{x^2 + y^2}{2}}\cos I
                - \sin I x\sqrt{1 - \frac{x^2 + y^2}{4}}}.
\end{align}
This is in agreement w/ Millholland \& Batygin 2019's Eq.~26. These have EOM
\begin{align*}
    \dot{y} = -\pd{H}{x} &=
        -x\p*{1 - \frac{x^2 + y^2}{2}} - \eta\p*{
            -x \cos I + \frac{x^2\sin I}{4}\p*{1 - \frac{x^2 + y^2}{4}}^{-1/2}
            - \sin I \sqrt{1 - \frac{x^2 + y^2}{4}}},\\
    \dot{x} = \pd{H}{y} &=
        y\p*{1 - \frac{x^2 + y^2}{2}} + \eta\p*{
            -y \cos I + \frac{xy\sin I}{4}\p*{1 - \frac{x^2 + y^2}{4}}^{-1/2}
            },\\
\end{align*}
The signs can be confirmed by considering that $\theta, x, y \approx 0$ in the
$\eta \to 0$ limit results in $\dot{\phi} < 0$, or at $\phi \approx 0, y \propto
\sin \phi \propto \phi$ and $\dot{y} < 0$. It bears noting that $x^2 + y^2 \in
[0, 4]$, so the EOM diverge as $\theta \to \pi$. This is one fewer singularity
than the naive $\p*{\cos\theta, \phi}$ covering, but full cartesian $\hat{s} =
\p*{s_x, s_y, s_z}$ still has no singularities.
